The Atala PRISM Protocol

### 2022-12-03 prism-backend/research/prism-protocol/did_atala.tex
Rewrote ideal functionality using global subroutines.

In this section we first give an ideal functionality for Public-Key
Infrastructures built atop of DIDs \IdealFPKIDID, and subsequently describe
Atala PRISM DID method \RealPKIDIDAtala, which is a protocol realizing
\IdealFPKIDID.
%
Before doing so, we introduce some base ideal functionalities upon which we
rely for \RealPKIDIDAtala. Namely, a simplified notion of append-only bulletin
boards, and digital signatures.

@@

\paragraph{Append-Only Bulletin Boards.} %
We adopt a simplified notion of append-only bulletin boards, given in
\cite{acc+20}. This is a local functionality, that defines an (probably too)
idealized version of a bulletin board, where no delays in the operations
occur \todo{We probably want to improve this to use a UC version that
  supports global functionalities}. For the sake of readability, we reproduce
the functionality in \figref{fig:faobb}.

\begin{figure}
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealFaobb(\cmd,\arg)$ from
        some party $P$ to run command \cmd, with arguments \arg,
        \IdealFaobb operates as follows.}
      \textrm{\IdealFaobb maintains a list \aobbl of bit strings.}      
    \end{minipage}
    \vspace*{0.5em}

    \begin{minipage}[t]{0.55\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Append} \land \arg = x$}{
        \pccomment{\fbox{Append data to \aobbl}} \\
        \aobbl \gets \aobbl || x \\
        \pcif P~\text{is corrupt}:~\ucsend~(\uccmd{Appended},x,P)~\text{to}
        ~\adv
      }
    \end{minipage}
    \hspace*{1.25em}
    \begin{minipage}[t]{0.45\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Retrieve} \land \arg =
        \emptyset$}{
        \pccomment{\fbox{Retrieve the contents of \aobbl}} \\
        \text{Output}~(\uccmd{Retrieved},\aobbl)~\text{to}~P
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal append-only bulletin board functionality, \IdealFaobb, from
    \cite{acc+20}.}
  \label{fig:faobb}
\end{figure}

\paragraph{Hashes.} %
\todo{\IdealFRO.}

@@

    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{KeyGen} \land \arg =
        \emptyset$}{
        \pccomment{\fbox{Key Generation}} \\
        \pcif \sid = (P,\ssid): \\
        \pcind \ucsend~(\uccmd{KeyGen},\sid)~\text{to}~\adv \\
        \pcind \ucrecv~(\uccmd{VerKey},\sid,v)~\text{from}~\adv \\
        \pcind \ucio{Output}~(\uccmd{VerKey},\sid,v)~\text{to}~P \\
        \pcind \text{Store}~(P,v)
      }
      \vspace*{0.25em}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Verif} \land
        \arg = (\sig,\msg,v')$}{
        \pccomment{\fbox{Verify a message}} \\
        \ucsend~(\uccmd{Verif},\sid,\msg,\sig,v)~\text{to}~\adv \\
        \ucrecv~(\uccmd{Verifd},\sid,\msg,\phi)~\text{from}~\adv \\
        \pcif v' = v \land (\msg,\sig,v,1) \in \SIG: f \gets 1 \\
        \pcelse \pcif v' = v \land P~\text{is not corrupt}~\land \\
        \pcind (\msg,\sig',v,1) \notin \SIG: \\
        \pcind \SIG \gets \SIG \cup \lbrace (\msg,\sig,v,0) \rbrace;
        f \gets 0 \\
        \pcelse \pcif (\msg,\sig,v',f') \in \SIG: f \gets f' \\
        \pcelse: \\
        \pcind \SIG \gets \SIG \cup \lbrace (\msg,\sig,v',\phi) \rbrace;~
        f \gets \phi \\
        \ucio{Output}~(\uccmd{Verifd},\sid,\msg,f)
      }
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Sign} \land \arg = \msg$}{
        \pccomment{\fbox{Sign a message}} \\
        \pcif \sid = (P,\ssid) \\
        \pcind \ucsend~(\uccmd{Sign},\sid,\msg)~\text{to}~\adv \\
        \pcind \ucrecv~(\uccmd{Signed},\sid,\msg,\sig)~\text{from}~
        \adv \\
        \pcind \pcif (\msg,\sig,v,0) \in \SIG: \\
        \pcind \pcind \ucio{Output}~(\uccmd{Error}) \\
        \pcind \pcelse: \\
        \pcind \pcind \SIG \gets \SIG \cup (\msg,\sig,v,1) \\
        \pcind \pcind \ucio{Output}~(\uccmd{Signed},\sid,\msg,\sig)
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal signature functionality, \IdealFSig, from \cite{canetti03}.}
  \label{fig:faobb}
\end{figure}

\subsection{Functionality \IdealFPKIDID}

\todo{Syntax using counter indexes doesn't seem a good idea (the simplification
  does not seem to work for proving security). Maybe change to arbitrary labels
  (and change current key label name to key type).}
\todo{Mention public delayed outputs.}
\todo{If the functionality tries to fetch some entry from a list, and the entry
  does not exist, then it aborts (assumed for updates).}

@@

Our expectation is that this simple approach allows to capture a large
subset of all possible DID method implementations. Note that the functional
restrictions we impose seem to be unavoidable. Namely, we do not specify how
the actual identifiers are computed; instead, we allow the adversary to pick
them. To update a DID, we only require such DID to exist and that the caller
is the party that created it, but otherwise the DID's contents can be set to
something completely new. Finally, deactivating labels within an existing DID
requires that these labels exist -- and, again, that the caller is the same
party who created the DID. Concretely, note that any further structure beyond
the label-type-value dictionary approach, is left as an implementation detail.
However, enforcing additional structure seems unavoidable in practice (e.g.,
requiring that all DIDs include keys of certain type). Since this definitely
has an impact on the behaviour of the ideal functionality, we parameterize it
with a policy \P, which can operate only on the received labels and key types,
and from which we only expect to return either $1$ to denote acceptance, or $0$
to denote rejection.

@@

\begin{figure}
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Upon receiving a call $\IdealFPKIDID^{\P}(\cmd,\sid,\arg)$
          from some party $P$ to run command \cmd, in session \sid, with arguments
          \arg, $\IdealFPKIDID^{\P}$ operates as follows.}
        \textrm{If the received \cmd and \arg pair does not meet any of
          the following options, the request is ignored.}
        \textrm{\P is a boolean predicate defining concrete conditions for
          acceptable DID creations, updates and deactivations.
        }
        \textrm{\todo{We assume that the functionality keeps track of the
            received requests.}}
      \end{minipage}
    }
    \vspace*{0.5em}

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]}$}{
          \pccomment{\fbox{DID Create operation}} \\
          \pcif \P(\lbrace (\lbl_i, \typ_i) \rbrace_{i\in[n]}) = 0:
          \text{abort} \\
          \ucsend~(\uccmd{Create}, \sid, \arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{CreateOk}, \sid, \did,
          \sval=\lbrace \val_i \rbrace_{i\in[n]}) \\
          \hspace*{2.8em}\text{from}~\adv \\
          \pcif (\cdot,\did,\cdot) \notin \DID: \\
          \pcind \DID \gets \DID \cup \lbrace (P,\did,\lbrace
          (\lbl_i,\typ_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
          \ucio{Output}~(\uccmd{Created},\sid,\did,\sval)
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
          \arg = did$}{
          \pccomment{\fbox{DID Read operation}} \\
          % \ucsend~(\uccmd{Read},\sid,\arg)~\text{to}~\adv \\
          % \ucrecv~(\uccmd{ReadOk},\sid)~\text{from}~\adv \\
          \pcif (\cdot,did,\sval) \in \DID: \\
          \pcind \ucio{Output}~(\uccmd{Read},\sid,(did,\sval))~\text{to}~P \\
          \pcelse: \\
          \pcind \ucio{Output}~(\uccmd{Read},\sid,\bot)~\text{to}~P    
        }    
      \end{minipage}
    }
    \vspace*{0.25em} \\
    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        % \arg = (did, \sval= \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]})\$}{
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (did, \sval_d,\sval_a)$}{
          \pccomment{\fbox{DID Update operation}} \\
          \lbrace (\lbl_i^a,\typ_i^a) \rbrace_{i\in[n^a]} \gets \sval_a \\
          \lbrace \lbl_i^d \rbrace_{i\in[n^d]} \gets \sval_d \\
          \pcif (P,\did,\sval_0) \notin \DID: \text{abort} \\
          % \sval_{del} \gets \sval_0 \setminus \sval;
          % \sval_{add} \gets \sval \setminus \sval_{del} \\          
          % n'' \gets |\sval_{add}| \\                    
          \ucsend~(\uccmd{Update}, \sid, \sval_a)~\text{to}~\adv \\
          \ucrecv~(\uccmd{UpdtOk},\sid,\lbrace \val^a_i \rbrace_{i\in[n^a]})~
          \text{from}~\adv \\
          \sval_1 \gets (\sval_0 \setminus \sval_d) \cup
          \lbrace (\lbl^a_i,\typ^a_i,\val^a_i) \rbrace_{i\in[n^a]} \\
          \pcif \P(\sval_1) = 0: \text{abort} \\
          \DID \gets \DID \setminus \lbrace (P,\did,\sval_0) \rbrace \cup
          \lbrace (P, \did, \sval_1) \rbrace \\
          \ucio{Output}~(\uccmd{Updated},\sid, (\did,\lbrace \val_a \rbrace_{i\in[n^a]}))
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (did, \lbrace \lbl_i \rbrace_{i\in[n]})$}{
          \pccomment{\fbox{DID Deactivate operation}} \\
          % \ucsend~(\uccmd{Deact}, \sid, \arg)~\text{to}~\adv \\
          % \ucrecv~(\uccmd{DeactOk},\sid)~\text{from}~\adv \\
          \pcif (P,did,\sval_0) \in \DID: \\
          \pcind \sval_1 \gets \sval_0 \setminus \lbrace \lbl_i
          \rbrace_{i\in[n]} \\
          \pcind \pcif \P(\sval_1) = 1: \\
          \pcind \pcind \DID \gets \DID \setminus \lbrace (P,did,\sval_0)
          \rbrace~\cup \\
          \hspace*{8em}\lbrace (P,did,\sval_1) \rbrace \\
          \ucio{Output}~(\uccmd{Deacted},\sid,\did)
        }
      \end{minipage}
    }
  \end{framed}
  \caption{Ideal DID-based PKI functionality, \IdealFPKIDID. We abuse set
    notation, and write $\lbrace (a_i,\dots) \rbrace_{i\in[n]} \setminus
    \lbrace (b_j,\dots) \rbrace_{j\in[m]}$ to mean subtracting from the first
    set the elements present in the second set, whenever $a_i = b_j$ for some
    $i,j$, even if the size of the $(a_i,\dots)$ and $(b_j,\dots)$ tuples do not
    match.}
  \label{fig:fpkidid}
\end{figure}

@@

We next define the Atala PRISM DID method, following the interfaces defined for
\IdealFPKIDID. We begin by describing the types of keys supported in Atala
PRISM, and then specify how sets of such keys can be managed through DIDs in
Atala.

@@

@@

Somehow related, it is of course not needed to fetch the contents of the whole
blockchain whenever we want to get an up to date DID Document. Instead, the
party running the operation can keep a local copy, and just fetch the new
operations since it last synced. However, again the given simplified notion of
a blockchain (via \IdealFaobb) eases notation -- and seems good enough for a
first modelling effort.

@@

\begin{theorem}[UC-security of \RealPKIDIDAtala]
  \label{thm:sec-didatala}
  \RealPKIDIDAtala UC-realizes $\IdealFPKIDID^{\P_{Atala}}$ in the $(\IdealFRO,
  \IdealFSig, \IdealFaobb)$-hybrid model.
\end{theorem}

>>>>

In this section we first give an ideal functionality for Public-Key
Infrastructures built atop of DIDs \IdealGPKIDID, and subsequently describe
Atala PRISM DID method \RealPKIDIDAtala, which is a protocol realizing
\IdealGPKIDID.
%
Before doing so, we introduce some base ideal functionalities upon which we
rely for \RealPKIDIDAtala. Namely, a simplified notion of append-only bulletin
boards, and digital signatures.

@@

\todo{Mention public delayed outputs.}
\todo{If the functionalities try to fetch some entry from a list, and the entry
  does not exist, then it aborts (assumed for updates).}

\paragraph{Hashes.} %
We model hashes as a global subroutine. No participant needs extra power (such
as programmability or observability), so we use a conventional random oracle --
called strict random oracle in \cite{cdg+18}. For self-containment, we reproduce
here the formulation.

\begin{figure}[ht]
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealGRO(\cmd,\sid,\arg)$ from some party
        $P$ to run command \cmd, in session \sid, with arguments \arg,
        \IdealGRO operates as follows.} \\
      \textrm{\IdealGRO keeps a list \hashlist of known hashed messages, and
        outputs hashes of length $l \gets l(n)$, for some fixed security
        parameter $n$.}
    \end{minipage}
    \vspace*{0.5em}

    \centering
    \procedure{$\pcif \cmd = \uccmd{HashQuery} \land \arg = \msg$}{%
      \pcif (\msg,h) \notin \hashlist:
      h \getr \bin^{l}; \hashlist \gets \hashlist \cup \lbrace (\msg,h) \rbrace \\
      \ucio{Output}~(\uccmd{HashConfirm},h)
    }
  \end{framed}
\end{figure}

\paragraph{Global clock.} %
Our ideal (global) functionality for a ledger relies on a global clock to model
delays in appending information. We rely on the ideal global clock functionality
given in \cite{kkk21}. In a nutshell, the functionality keeps track of the
parties in the system, and allows them to signal their intention to advance the
clock's time. When all parties agree on it, the clock's time is incremented. See
\figref{fig:gclock}.

\begin{figure}[ht]
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealGclock(\cmd,\arg)$ from
        some party $P$ to run command \cmd, with arguments \arg,
        \IdealGclock operates as follows.} \\
      \textrm{\IdealGclock maintains a variable $t$ with the current itme,
        a variable $T$ with the list of parties keeping track of the time,
        and a map $A$ which keeps the parties' (dis)agreements to advance the
        clock.}
    \end{minipage}
    \vspace*{0.5em}

    \begin{minipage}[t]{0.40\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Register} \land \arg = \emptyset$}{
        T \gets T \cup \set{P} \\
        \ucio{Output}~(\uccmd{Registered})~\text{to}~P \\
      }
      \procedure{$\pcif \cmd = \uccmd{Update} \land \arg = \emptyset$}{
        A(P) \gets \top \\
        \pcif \forall P \in T, A(P) = \top: \\
        \pcind t \gets t + 1; \forall P \in T, A(P) = \bot \\
        \pcind \ucio{Send}~(\uccmd{Updated})~\text{to}~P        
      }
    \end{minipage}
    \hspace*{1.25em}
    \begin{minipage}[t]{0.60\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Deregister} \land \arg = \emptyset$}{
        T \gets T \setminus \set{P} \\
        \text{Output}~(\uccmd{DeregisteredOk})~\text{to}~P \\
      }
      \procedure{$\pcif \cmd = \uccmd{Read} \land \arg = \emptyset$}{
        \text{Output}~(\uccmd{ReadOk},t)~\text{to}~P
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal subroutine for a global clock, from \cite{kkk21}.}
  \label{fig:gclock}
\end{figure}

\paragraph{$\delta$-delay Ledger.} %
We adopt a ledger notion given in \cite{kkk21}. This is global functionality
that defines an ideal version of a ledger as a shared subroutine, which models
the append-only property of ledgers, and the fact that changes take some delay
until they are applied to the ledger.
%
This functionality accepts sending individual transactions to be added to the
ledger state, fetching (reading) the state of the ledger, and extending the
ledger by appending (blocks of) transactions to it. Interestingly, it gives the
adversary control over when to advance the status of each party, as long as
these advancements ensure the common prefix rule -- where $\ldgState' \prec
\ldgState$ means that $\ldgState'$ is a prefix of \ldgState. It also relies
on yet another global functionality, the global clock, to timestamp incoming
transactions. If a transaction is not appended to an honest party's view before
the given $\delta$ delay, then the functionality aborts.
%
For the sake of readability, we reproduce the functionality in
\figref{fig:gledger}.

\begin{figure}[ht]
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealGdledger(\cmd,\arg)$ from
        some party $P$ to run command \cmd, with arguments \arg,
        \IdealGdledger operates as follows.} \\
      \textrm{\IdealGdledger maintains a mapping \ldgMap which associates
        parties with their view of the ledger.
        \ldgUtxo is a multiset of unconfirmed transactions.
        \ldgState is the authoritative ledger state
        All mappings and (multi-)sets are initialised to empty values.
      }
    \end{minipage}
    \vspace*{0.5em}

    \begin{minipage}[t]{0.40\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Submit} \land \arg = \ldgTx$}{
        \ucio{Send}~(\uccmd{Read})~\text{to}~\IdealGclock \\
        \ucio{Recv}~(\uccmd{ReadOk},t)~\text{from}~\IdealGclock \\
        \ldgUtxo \gets \ldgUtxo \cup \set{(\ldgTx,t)} \\ 
        \ucio{Output}~(\uccmd{Transaction},\ldgTx,t)~\text{to}~\adv \\
      }
      \procedure{$\pcif \cmd = \uccmd{Extend} \land \arg = \ldgState'$}{
        \pcif P = \adv \land \ldgState' \subseteq \ldgUtxo: \\
        \pcind \ldgUtxo \gets \ldgUtxo \setminus \ldgState' \\
        \pcind \ldgState \gets \ldgState || \ldgState'
      }
    \end{minipage}
    \hspace*{1.25em}
    \begin{minipage}[t]{0.60\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Read} \land \arg =
        \emptyset$}{
        % \pcif P = \adv: data \gets \ldgState \\
        % \pcelse: data \gets \ldgMap(P) \\
        \ucio{Send}~(\uccmd{Read})~\text{to}~\IdealGclock \\
        \ucio{Recv}~(\uccmd{ReadOk},t)~\text{from}~\IdealGclock \\
        \pcif \exists (\ldgTx,t') \in \ldgUtxo~\suchthat~|t-t'| > \delta:
        \pcreturn \bot \\
        \pcelse \pcif \exists (\ldgTx,t') \in \ldgState~\suchthat~ |t-t'|
        > \delta~\land \\
        \pcind \exists P' \in \ldgHonest~\suchthat~(\ldgTx,t') \notin \ldgMap(P'):
        \pcreturn \bot \\        
        \text{Output}~(\uccmd{ReadOk}, \ldgHead(\ldgMap(P)))~\text{to}~P \\
      }
      \procedure{$\pcif \cmd = \uccmd{Advance} \land \arg = (P',\ldgState')$}{
        \pcif P = \adv \land \ldgMap(P') \prec \ldgState' \prec \ldgState:
        \ldgMap(P') \gets \ldgState'
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal subroutine for a global $\delta$-delay ledger, from
    \cite{kkk21}.}
  \label{fig:gledger}
\end{figure}

@@

    \begin{minipage}[t]{0.5\textwidth}
      \procedure{$\pcif \cmd = \uccmd{KeyGen} \land \arg =
        \emptyset$}{
        \pccomment{\fbox{Key Generation}} \\
        \pcif \sid = (P,\ssid): \\
        \pcind \ucsend~(\uccmd{KeyGen},\sid)~\text{to}~\adv \\
        \pcind \ucrecv~(\uccmd{VerKey},\sid,v)~\text{from}~\adv \\
        \pcind \ucio{Output}~(\uccmd{VerKey},\sid,v)~\text{to}~P \\
        \pcind \text{Store}~(P,v)
      }
      \vspace*{0.25em}
      \procedure{$\pcif \cmd = \uccmd{Verif} \land
        \arg = (\sig,\msg,v')$}{
        \pccomment{\fbox{Verify a message}} \\
        \ucsend~(\uccmd{Verif},\sid,\msg,\sig,v)~\text{to}~\adv \\
        \ucrecv~(\uccmd{Verifd},\sid,\msg,\phi)~\text{from}~\adv \\
        \pcif v' = v \land (\msg,\sig,v,1) \in \SIG: f \gets 1 \\
        \pcelse \pcif v' = v \land P~\text{is not corrupt}~\land \\
        \pcind (\msg,\sig',v,1) \notin \SIG: \\
        \pcind \SIG \gets \SIG \cup \lbrace (\msg,\sig,v,0) \rbrace;
        f \gets 0 \\
        \pcelse \pcif (\msg,\sig,v',f') \in \SIG: f \gets f' \\
        \pcelse: \\
        \pcind \SIG \gets \SIG \cup \lbrace (\msg,\sig,v',\phi) \rbrace;~
        f \gets \phi \\
        \ucio{Output}~(\uccmd{Verifd},\sid,\msg,f)
      }
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
      \procedure{$\pcif \cmd = \uccmd{Sign} \land \arg = \msg$}{
        \pccomment{\fbox{Sign a message}} \\
        \pcif \sid = (P,\ssid) \\
        \pcind \ucsend~(\uccmd{Sign},\sid,\msg)~\text{to}~\adv \\
        \pcind \ucrecv~(\uccmd{Signed},\sid,\msg,\sig)~\text{from}~
        \adv \\
        \pcind \pcif (\msg,\sig,v,0) \in \SIG: \\
        \pcind \pcind \ucio{Output}~(\uccmd{Error}) \\
        \pcind \pcelse: \\
        \pcind \pcind \SIG \gets \SIG \cup (\msg,\sig,v,1) \\
        \pcind \pcind \ucio{Output}~(\uccmd{Signed},\sid,\msg,\sig)
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal signature functionality, \IdealFSig, from \cite{canetti03}.}
  \label{fig:fsig}
\end{figure}

\subsection{Global Functionality \IdealGPKIDID}

@@

Our expectation is that this simple approach allows to capture a large
subset of all possible DID method implementations. Note that the functional
restrictions we impose seem to be unavoidable. Namely, we do not specify how
the actual identifiers are computed; instead, we allow the adversary to pick
them. To update a DID, we only require such DID to exist and that the caller
is the party that created it, but otherwise the DID's contents can be set to
something completely new. Finally, deactivating labels within an existing DID
requires that these labels exist -- and, again, that the caller is the same
party who created the DID. Concretely, note that any further structure beyond
the label-type-value dictionary approach, is left as an implementation detail.
However, enforcing additional structure seems unavoidable in practice (e.g.,
requiring that all DIDs include keys of certain type). Since this definitely
has an impact on the behaviour of the ideal functionality, we parameterize it
with a policy \P, which can be used as a black box: the functionality invokes
\P with a set of labels and key types, and receives either $1$ to denote
acceptance, or $0$ to denote rejection. However, it is not realistic to assume
that, even corrupted parties will enforce such policies. Thus, our approach is
to defer policy checking in the ideal functionality until after having
interacted with the simulator \Sim. Intuitively then, if the output of the
ideal functionality along with \Sim is indistinguishable to that of the real
protocol with any real world adversary, then the policy is correctly enforced
by the real protocol. Looking ahead, the real protocol in turn ``delegates''
this policy-checking to an ideal append-only bulletin board. \todo{This captures
  the idea, but refine and re-structure.}

@@

\begin{figure}
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Upon receiving a call $\IdealGPKIDID^{\P}(\cmd,\sid,\arg)$
          from some party $P$ to run command \cmd, in session \sid, with arguments
          \arg, $\IdealGPKIDID^{\P}$ operates as follows.}
        \textrm{If the received \cmd and \arg pair does not meet any of
          the following options, the request is ignored.}
        \textrm{\P is a boolean predicate defining concrete conditions for
          acceptable DID creations, updates and deactivations.}
        \textrm{\didOP is a list of submitted (yet, possibly unconfirmed) DID
          operations, and \DID is the authoritative state of published DIDs.}
      \end{minipage}
    }
    \vspace*{0.5em}

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \setind{(\lbl_i,\typ_i)}{i\in[n]}$}{ 
          \DID \gets \DIDReadLdg \\
          \pcif (\did,\cdot) \in \DID \lor
          \P(\setind{(\lbl_i,\typ_i)}{i\in[n]}) = 0: \\
          \pcind \pcreturn \bot \\
          \ucsend~(\uccmd{Create}, \sid, \arg)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{CreateOk}, \sid, \did,
          \sval=\setind{\val_i}{i\in[n]}) \\
          \hspace*{2.7em}\text{from}~\Sim \\          
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \didOp \gets (\did,\setind{(\lbl_i,\typ_i,\val_i)}{i\in[n]}) \\
          \didOP \gets \didOP \cup \set{(P,\tau,\didOp)} \\
          \ucsend~(\uccmd{Submit}, \didOp)~\text{to}~\IdealGdledger \\
          \ucio{Output}~(\uccmd{ProcessedC},\sid,\did,\sval)
          % \pcind \DID \gets \DID \cup \lbrace (P,\did,\lbrace
          % (\lbl_i,\typ_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
          % \ucio{Output}~(\uccmd{Created},\sid,\did,\sval,t)
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
          \arg = \did$}{
          \DID \gets \DIDReadLdg \\
          \pcif (\did,\sval) \notin \DID \lor \P(\sval) = 0: \\
          \pcind \pcreturn \bot \\
          \ucio{Output}~(\uccmd{ReadOk},\sid,\did,\sval)~\text{to}~P
        }    
      \end{minipage}
    }
    \vspace*{0.25em} \\
    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        % \arg = (did, \sval= \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]})\$}{
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (did, \sval_d,\sval_a)$}{
          \DID \gets \DIDReadLdg \\
          \pcif (\did,\sval_0) \notin \DID: \pcreturn \bot \\
          \pcif \P((\sval_0 \setminus \sval_d) \cup \sval_a) = 0:
          \pcreturn \bot \\
          \ucsend~(\uccmd{Update}, \sid,\did,\sval_d,\sval_a)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{UpdOk},\sid,\did,\lbrace \val^a_i \rbrace_{i\in[n^a]})~
          \text{from}~\Sim \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \lbrace (\lbl_i^a,\typ_i^a) \rbrace_{i\in[n^a]} \gets \sval_a \\
          %\lbrace \lbl_i^d \rbrace_{i\in[n^d]} \gets \sval_d \\          
          \sval_1 \gets (\sval_0 \setminus \sval_d) \cup
          \setind{(\lbl^a_i,\typ^a_i,\val^a_i)}{i\in[n^a]} \\
          \didOp \gets (\did,\sval_1);~
          \didOP \gets \didOP \cup \set{(P,\tau,\didOp)} \\
          \ucsend~(\uccmd{Submit}, \didOp)~\text{to}~\IdealGdledger \\
          \ucio{Output}~(\uccmd{ProcessedU},\sid,\did,\didOp)
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (did, \lbrace \lbl_i \rbrace_{i\in[n]})$}{
          \DID \gets \DIDReadLdg \\
          \pcif (P,did,\sval_0) \notin \DID: \pcreturn \bot \\
          \sval_1 \gets \sval_0 \setminus \lbrace \lbl_i \rbrace_{i\in[n]} \\
          \pcif \P(\sval_1) = 0: \pcreturn \bot \\
          \ucsend~(\uccmd{Deact}, \sid, \arg)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{DeactOk},\sid)~\text{from}~\Sim \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\          
          \didOp \gets (\did,\sval_1);
          \didOP \gets \didOP \cup \set{(P,\tau,\didOp)} \\
          \ucsend~(\uccmd{Submit},\didOp)~\text{to}~\IdealGdledger \\          
          \ucio{Output}~(\uccmd{ProcessedD},\sid,\did,\didOp)
        }
      \end{minipage}
    }
    \vspace*{0.25em} \\
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \centering
        \procedure[linenumbering]{$\DIDReadLdg$}{
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGdledger~\text{and recv}~
          (\uccmd{ReadOk},\DID)~\text{back} \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock ~\text{and recv}~
          (\uccmd{ReadOk},\tau)~\text{back} \\
          \pcif \exists (\cdot,\tau',\didOp) \in \didOP~\suchthat~
          |\tau - \tau'| > \delta \land \P(\didOp) = 1 \land \didOp \notin \DID:
          \pcreturn \bot \\
          \pcreturn \DID
        }
      \end{minipage}
    }
  \end{framed}
  \caption{Ideal DID-based PKI functionality, \IdealGPKIDID. We abuse set
    notation, and write $\lbrace (a_i,\dots) \rbrace_{i\in[n]} \setminus
    \lbrace (b_j,\dots) \rbrace_{j\in[m]}$ to mean subtracting from the first
    set the elements present in the second set, whenever $a_i = b_j$ for some
    $i,j$, even if the size of the $(a_i,\dots)$ and $(b_j,\dots)$ tuples do not
    match.}
  \label{fig:fpkidid}
\end{figure}

@@

We next define the Atala PRISM DID method, following the interfaces defined for
\IdealGPKIDID. We begin by describing the types of keys supported in Atala
PRISM, and then specify how sets of such keys can be managed through DIDs in
Atala.

@@

\paragraph{$\P_{Atala}$ policy.} %
Before proceeding with the specififying the DID operations in Atala, we define
its policy on the keys' labels and types.

\begin{figure}[ht]
  \centering
  \procedure{$\P_{Atala}(\lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]})$}{    
    \pcif \exists i\in[n]~\suchthat~\typ_i \notin [\MasterKey,\IssueKey,
    \CommKey,\AuthKey]: \pcreturn 0~\pccomment{All types must be known} \\   
    \pcif \not\exists i\in[n]~\suchthat~\typ_i = \MasterKey: \pcreturn 0~
    \pccomment{There must be at least one master key} \\
    \pcif \exists i, j \in[n], i \neq j~\suchthat~\lbl_i = \lbl_j: \pcreturn 0~
    \pccomment{No duplicate labels} \\
    \pcreturn 1
  }
  \caption{Policy $\P_{Atala}$ that all DID documents must meet in Atala.}
\end{figure}

@@

Somehow related, it is of course not needed to fetch the contents of the whole
blockchain whenever we want to get an up to date DID Document. Instead, the
party running the operation can keep a local copy, and just fetch the new
operations since it last synced. However, again the given simplified notion of
a blockchain (via \IdealGdledger) eases notation -- and seems good enough for a
first modelling effort.

@@

\todo{Do we mention somewhere that {\sid}s have structure $(P,\sid')$? This is
  used for \Sim to know which party is generating which keys.}

\begin{theorem}[UC-security of \RealPKIDIDAtala]
  \label{thm:sec-didatala}
  \RealPKIDIDAtala UC-realizes $\IdealGPKIDID^{\P_{Atala}}$ in the $(\IdealGRO,
  \IdealFSig, \IdealGdledger)$-hybrid model.
\end{theorem}

\begin{proof}
  We show that \RealPKIDIDAtala UC-realizes \IdealGPKIDID, when configured
  with policy $\P_{Atala}$. For this, we give a simulator \Sim such that, for
  any environment \Env, $\Exec_{\RealPKIDIDAtala\adv,\Env} \approx
  \Exec_{\phiDID,\Sim,\Env}$,
  where \phiDID is a protocol composed with the same parties as
  \RealPKIDIDAtala but where each party is simply a dummy party that forwards
  every message from \Env to the ideal functionality \IdealGPKIDID, and every
  response from \IdealGPKIDID back to \Env. The simulator \Sim runs internally a
  copy of the real-world adversary \adv, as well as the ideal functionalities
  \IdealGRO, \IdealFSig and \IdealGdledger, and any involved party.
